---
title: "Sobol Sensitivity Analysis Assignment"
author: "Your Name"
date: today
format: 
  html:
    code-fold: false
    toc: true
    theme: cosmo
  pdf:
    toc: true
---

# Part I: Paper Review

## Uranium in Groundwater Model

In the uranium paper, the sensitivity analysis revealed that uranium concentration accounts for 69-78% of output variance, suggesting that water quality monitoring should be the focus of management resources. Also, since body weight contributes <1% to risk variance, assessments should use standard default values rather than conducting demographic surveys of the exposed population. Resources should instead be concentrated on monitoring groundwater uranium levels, particularly in the 17% of locations where levels exceed WHO guidelines. We can also note that uncertainty in uranium concentration will propagate most strongly to risk estimates, so investment in high quality concentration analysis will have the highest payoff for improving risk assessment accuracy.

# Part II: Atmospheric Conductance Sensitivity Analysis

## Setup
```{r}
#| warning: false
#| message: false

library(sensitivity)
library(tidyverse)
library(purrr)

source(here::here("R/Catm.R"))

# Generate parameter sets using sobolSalt

# Set sample size
np <- 1000

# Generate first set of parameters (X1)

# Windspeed: higher and more variable than class (mean=300, sd=50)
# Height: shorter vegetation (3.5-5.5m vs 9.5-10.5m in class)
k_o <- rnorm(mean = 0.1, sd = 0.1 * 0.01, n = np)  # 1% SD
k_d <- rnorm(mean = 0.7, sd = 0.7 * 0.01, n = np)  # 1% SD  
v <- rnorm(mean = 300, sd = 50, n = np)            # Higher, more variable
height <- runif(min = 3.5, max = 5.5, n = np)      # Shorter vegetation

X1 <- cbind.data.frame(k_o, k_d, v, height)

# Generate second set of parameters (X2) 
k_o <- rnorm(mean = 0.1, sd = 0.1 * 0.01, n = np)
k_d <- rnorm(mean = 0.7, sd = 0.7 * 0.01, n = np)
v <- rnorm(mean = 300, sd = 50, n = np)
height <- runif(min = 3.5, max = 5.5, n = np)

X2 <- cbind.data.frame(k_o, k_d, v, height)

# Create Sobol object
sens_Catm_Sobol <- sobolSalt(model = NULL, X1, X2, nboot = 100)
```

```{r}
# Run model for all parameter sets

parms <- as.data.frame(sens_Catm_Sobol$X)
colnames(parms) <- c("k_o", "k_d", "v", "height")

# Run the model for all parameter combinations
res <- pmap_dbl(parms, Catm)

# Give results to sobol object
sens_Catm_Sobol <- sensitivity::tell(sens_Catm_Sobol, res, res.names = "Catm")
```

```{r}
# Calculate and display sobol indices

# First-order effects (main effects)
row.names(sens_Catm_Sobol$S) <- colnames(parms)
print("First-order sensitivity indices:")
print(sens_Catm_Sobol$S)

# Total effects (includes interactions)
row.names(sens_Catm_Sobol$T) <- colnames(parms)
print("Total effect sensitivity indices:")
print(sens_Catm_Sobol$T)

# Print full summary
print(sens_Catm_Sobol)
```

```{r}
# Combine parameters and results
both <- cbind.data.frame(parms, Catm = sens_Catm_Sobol$y)

# Plot 1: Histogram showing uncertainty in conductance
ggplot(both, aes(x = Catm)) +
  geom_histogram(bins = 50, fill = "skyblue", color = "black") +
  geom_vline(xintercept = mean(both$Catm), col = "red", linewidth = 1) +
  labs(title = "Uncertainty in Atmospheric Conductance",
       subtitle = "High windspeed, short vegetation setting",
       x = "Conductance (mm/s)",
       y = "Frequency") +
  theme_minimal(base_size = 12)
```

```{r}
# Identify the top 2 most sensitive parameters by total effect
sensitivity_ranking <- sens_Catm_Sobol$T %>%
  as.data.frame() %>%
  mutate(param = row.names(.)) %>%
  arrange(desc(original))

most_sensitive <- sensitivity_ranking$param[1]
second_sensitive <- sensitivity_ranking$param[2]

print(paste("Most sensitive parameter:", most_sensitive))
print(paste("Second most sensitive parameter:", second_sensitive))
```

```{r}
# Plot conductance vs windspeed, colored by 2nd most sensitive parameter
ggplot(both, aes(x = v, y = Catm, color = .data[[second_sensitive]])) +
  geom_point(alpha = 0.6) +
  scale_color_viridis_c() +
  labs(title = "Atmospheric Conductance vs Windspeed",
       subtitle = paste("Colored by", second_sensitive, "(2nd most sensitive)"),
       x = "Windspeed (m/s)",
       y = "Conductance (mm/s)",
       color = second_sensitive) +
  theme_minimal(base_size = 12)
```

```{r}
# Calculate interaction effects (Total - First order)
interactions <- sens_Catm_Sobol$T$original - sens_Catm_Sobol$S$original
names(interactions) <- row.names(sens_Catm_Sobol$S)

print("Interaction effects (Total - First order):")
print(interactions)
```

