---
title: "Sobol Sensitivity Analysis Assignment"
author: "Emily Miller"
date: today
format: pdf
---

# Part I: Paper Review

## Uranium in Groundwater Model

In the uranium paper, the sensitivity analysis revealed that uranium concentration accounts for 69-78% of output variance, suggesting that water quality monitoring should be the focus of management resources. Also, since body weight contributes <1% to risk variance, assessments should use standard default values rather than conducting demographic surveys of the exposed population. Resources should instead be concentrated on monitoring groundwater uranium levels, particularly in the 17% of locations where levels exceed WHO guidelines. We can also note that uncertainty in uranium concentration will propagate most strongly to risk estimates, so investment in high quality concentration analysis will have the highest payoff for improving risk assessment accuracy.

# Part II: Atmospheric Conductance Sensitivity Analysis

## Setup
```{r}
#| warning: false
#| message: false

library(sensitivity)
library(tidyverse)
library(purrr)

source(here::here("R/Catm.R"))

# Generate parameter sets using sobolSalt

# Set sample size
np <- 1000

# Generate first set of parameters (X1)

# Windspeed: higher and more variable than class (mean=300, sd=50)
# Height: shorter vegetation (3.5-5.5m vs 9.5-10.5m in class)
k_o <- rnorm(mean = 0.1, sd = 0.1 * 0.01, n = np)  # 1% SD
k_d <- rnorm(mean = 0.7, sd = 0.7 * 0.01, n = np)  # 1% SD
v <- rnorm(mean = 300, sd = 50, n = np)            # Higher, more variable
height <- runif(min = 3.5, max = 5.5, n = np)      # Shorter vegetation

X1 <- cbind.data.frame(k_o, k_d, v, height)

# Generate second set of parameters (X2) 
k_o <- rnorm(mean = 0.1, sd = 0.1 * 0.01, n = np)
k_d <- rnorm(mean = 0.7, sd = 0.7 * 0.01, n = np)
v <- rnorm(mean = 300, sd = 50, n = np)
height <- runif(min = 3.5, max = 5.5, n = np)

X2 <- cbind.data.frame(k_o, k_d, v, height)

# Create Sobol object
sens_Catm_Sobol <- sobolSalt(model = NULL, X1, X2, nboot = 100)
```

```{r}
# Run model for all parameter sets

parms <- as.data.frame(sens_Catm_Sobol$X)
colnames(parms) <- c("k_o", "k_d", "v", "height")

# Run the model for all parameter combinations
res <- pmap_dbl(parms, Catm)

# Give results to sobol object
sens_Catm_Sobol <- sensitivity::tell(sens_Catm_Sobol, res, res.names = "Catm")
```

```{r}
# Calculate and display sobol indices

# First-order effects (main effects)
row.names(sens_Catm_Sobol$S) <- colnames(parms)
print("First-order sensitivity indices:")
print(sens_Catm_Sobol$S)

# Total effects (includes interactions)
row.names(sens_Catm_Sobol$T) <- colnames(parms)
print("Total effect sensitivity indices:")
print(sens_Catm_Sobol$T)

# Print full summary
print(sens_Catm_Sobol)
```

```{r}
# Combine parameters and results
both <- cbind.data.frame(parms, Catm = sens_Catm_Sobol$y)

# Histogram showing uncertainty in conductance
ggplot(both, aes(x = Catm)) +
  geom_histogram(bins = 50, fill = "skyblue", color = "black") +
  geom_vline(xintercept = mean(both$Catm), col = "red", linewidth = 1) +
  labs(title = "Uncertainty in Atmospheric Conductance",
       subtitle = "High windspeed, short vegetation setting",
       x = "Conductance (mm/s)",
       y = "Frequency") +
  theme_minimal(base_size = 12)
```

```{r}
# Plot conductance vs windspeed, colored by height (2nd most sensitive parameter) 
ggplot(both, aes(x = v, y = Catm, color = height)) +
  geom_point(alpha = 0.6) +
  scale_color_viridis_c() +
  labs(title = "Atmospheric Conductance vs Windspeed",
       subtitle = "Colored by height (2nd most sensitive)",
       x = "Windspeed (m/s)",
       y = "Conductance (mm/s)",
       color = "Height (m)") +
  theme_minimal(base_size = 12)
```

```{r}
# Calculate interaction effects (Total - First order)
interactions <- sens_Catm_Sobol$T$original - sens_Catm_Sobol$S$original
names(interactions) <- row.names(sens_Catm_Sobol$S)

print("Interaction effects (Total - First order):")
print(interactions)
```

## Discussion: Comparison with Class Setting

This sensitivity analysis shows windspeed (v) dominates conductance uncertainty, with a first-order index of 0.82 (82%), far higher than the class setting where v contributed only 18%. This increased sensitivity results from higher v variability (SD = 50 vs 30 m/s) and mean (300 vs 250 m/s). k_d was the most sensitive parameter in class (55%), barely but contributes here (~5%), indicating that in shorter vegetation with higher windspeed, the direct effect of wind overwhelms turbulence parameters. Vegetation height is the second most sensitive parameter (first-order index = 0.22) in this setting, whereas it was essentially insensitive in the tall forest class setting (<1%). Interaction effects are fairly small (all < 0.06), indicating parameters act relatively independently.
